---
title: "Reproducible-research-project-2-Markdown"
author: "Ing. Greg Bennett"
date: "August 30, 2017"
output:
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data

This is an **R Markdown** document contains the Peer-graded Assignment: Course Project 2. 
Step 1
You will need to download & save in a specific folder, see: <https://github.com/gbg184/reproducible-research-project-2-Markdown>.

```{r read}
library(readr)
df <- read_csv("C:/Users/gbennett/Dropbox/Data Scientists/5.Reproducible_Research/project/repdata_data_StormData.csv")
```

## Process the data

To evaluate the health impact, the total fatalities and the total injuries for each event type (EVTYPE) are calculated. The codes for this calculation are:
```{r pressure, echo=FALSE}
df.fatalities <- df %>% select(EVTYPE, FATALITIES) %>% group_by(EVTYPE) %>% summarise(total.fatalities = sum(FATALITIES)) %>% arrange(-total.fatalities)
head(df.fatalities, 10)

df.injuries <- df %>% select(EVTYPE, INJURIES) %>% group_by(EVTYPE) %>% summarise(total.injuries = sum(INJURIES)) %>% arrange(-total.injuries)
head(df.injuries, 10)

df.damage <- df %>% select(EVTYPE, PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP)
```
Economic impact:
```{r Impact, echo=FALSE}
 df.damage$Prop.Multiplier <- x$Multiplier[match(df.damage$PROPDMGEXP,x$Symbol)]
 df.damage$Prop.Multiplier <- x$Multiplier[match(df.damage$CROPDMGEXP,x$Symbol)]
 
```

## Results
Health Impact
The **top 10 events** with the highest total fatalities and injuries are shown graphically.

```{r result}
g <- ggplot(df.fatalities[1:10,], aes(x=reorder(EVTYPE, -total.fatalities), y=total.fatalities))+geom_bar(stat="identity") + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+ggtitle("Top 10 Events with Highest Total Fatalities") +labs(x="EVENT TYPE", y="Total Fatalities")
g <- ggplot(df.injuries[1:10,], aes(x=reorder(EVTYPE, -total.injuries), y=total.injuries))+geom_bar(stat="identity") + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+ggtitle("Top 10 Events with Highest Total Injuries") +labs(x="EVENT TYPE", y="Total Injuries")
g

```
